---
title: "Covid91 vaccine study Final2023"
author: "Steven Garcia"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ There is no difference between the affect of the covid921 to male population.

$H_A$ There is a difference between the affect of the covid921 test trials on male population.

```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex == 'M')
```

## Methods

Males is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```

This bar graph shows among the males who received the placebo had a higher percentage of getting Covid19. Compared to males who received the vaccine, the people had a significant.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

According to this table the percentage of males that were infected with the placebo were 2 times more likely to become infected. The drug was a whole lot more effective than those who did not get vaccine. Those who did not receive the drug (none) had a 50% chance of becoming infected.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0.000000000232, which is below 0.05, so we reject the null hypothesis. This indicates that there is a difference among those who receive the drug and those who do not. 

You are almost 2.24 times more likely to be protected from getting COVID vaccine compared to those who didn't.

# Females

$H_0$ There is no difference between the affect of the covid921 to female population.

$H_A$ There is a difference between the affect of the covid921 test trials on female population.

```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex == 'F')
```

## Methods

Females is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

This bar graph shows among the females who received the placebo had a higher percentage of getting Covid19. Compared to Females who received the vaccine, those females had a significant. It was a 50-50 chance for those who never caught Covid.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

According to this table the percentage of Females that were infected with the placebo were 1-1/2 times more likely to become infected. The drug was a whole lot more effective than those who did not get vaccine. Those who did not receive the drug (none) had a 50% chance of becoming infected. Unfortunately, they had a high chance of attracting the virus at 55% chance overall. 


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0.00002732, which is below 0.05, so we reject the null hypothesis. This indicates that there is a difference among those who receive the drug and those who do not. 

You are almost 1.5 times more likely to be protected from getting COVID vaccine compared to those who didn't. 

# LGBTQ

$H_0$ There is no difference between the affect of the covid921 to LGBTQ population.

$H_A$ There is a difference between the affect of the covid921 test trials on LGBTQ population.

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ == "gay")
```

## Methods

LGBTQ is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data= LGBTQ)
barchartGC(~infected + treatment,data= LGBTQ, type="percent")
```

This bar graph shows among the LGBTQ who received the vaccine had a higher percentage of getting Covid19. Compared to LGBTQ who received the placebo, those LGBTQ had a lower chance of catching covid19. Those who never caught covid had little change. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

According to this table the percentage of LGBTQ that were infected with the vaccine had a 84% chance of more likely to become infected. The drug was a whole a lot less effective than those who did not get vaccine. Those who receive the placebo drug (none) had a greater chance of not becoming infected.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0.0002732, which is below 0.05, so we reject the null hypothesis. This indicates that there is a difference among those who receive the drug and those who do not. 

You are almost 1.5 times more likely to get COVID with the vaccine compared to those who didn't.  

# Druggies

$H_0$ There is no difference between the affect of the covid921 to druggies population.

$H_A$ There is a difference between the affect of the covid921 test trials on druggies population.

```{r}
library(dplyr)
DrugUsers = filter(FinalData, FinalData$DrugUser == 'yes')
```

## Methods

Druggies is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")
```

This bar graph shows among the DrugUsers who received the vaccine had a higher percentage of getting Covid19. Compared to DrugUsers who received the placebo, those druggies had a lower chance of catching covid19. Those who never caught covid had little change. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```

According to this table the percentage of DrugUsers that were infected with the placebo were 8 times less  likely to become infected. The drug was a whole lot less effective than those who got the placebo. Those who did not receive the drug (none) had a 5-50% chance of becoming infected.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0.0002732, which is below 0.05, so we reject the null hypothesis. This indicates that there is a difference among those who receive the drug and those who do not. 

You are almost 1.5 times more likely to catch COVID if you were to get the vaccine. 


# Overall Results and Conclusions

Does the vaccine prove to be effective in different effectiveness across a variety of sub-population? According to the males and females chisq test which resulted in a P value less than 0.05. Therefore we reject the null hypothesis. The vaccine is effective in preventing COVID921 for those two groups. Conversely, according to the LGBTQ and DrugUsers chisq test which resulted in a P value less than 0.05. Therefore, we reject the null hypothesis. The vaccine is less effective when it comes to preventing COVID921. The last two groups are 8 times more likely to contract COVID921 when given the vaccine. Those who received the placebo are significantly less likely to contract COVID921. 
