---
title: "Covid91 vaccine study Final2023"
author: "Steven Garcia"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ There is no difference between the affect of the covid921 to male population.

$H_A$ There is a difference between the affect of the covid921 test trials on male population.

```{r}
library(dplyr)
Males = filter(FinalData, FinalData$sex == 'M')
```

## Methods

Males is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```

This bar graph shows among the males who received the placebo had a higher percentage of getting Covid19. Compared to males who received the vaccine, the people had a significant.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

## Graphical Descriptive Result

catergorical~categorical

table1 <- xtabs(~class + Sex, data=druggiesData)

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Females

$H_0$ There is no difference between the affect of the covid921 to female population.

$H_A$ There is a difference between the affect of the covid921 test trials on female population.

```{r}
library(dplyr)
Females = filter(FinalData, FinalData$sex == 'F')
```

## Methods

Females is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

This bar graph shows among the females who received the placebo had a higher percentage of getting Covid19. Compared to males who received the vaccine, those females had a significant.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```



```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ == 'gay')
head(LGBTQ)
```


table1 <- xtabs(~class + Sex, data=davisdata)

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# LGBTQ

$H_0$ There is no difference between the affect of the covid921 to LGBTQ population.

$H_A$ There is a difference between the affect of the covid921 test trials on LGBTQ population.

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ == "gay")
```

## Methods

LGBTQ is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

This bar graph shows among the LGBTQ who received the vaccineo had a higher percentage of getting Covid19. Compared to LGBTQ who received the placebo, those LGBTQ had a lower chance of catching covid19. Those who never caught covid had little change. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```



table1 <- xtabs(~class + Sex, data=davisdata)

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Druggies

$H_0$ There is no difference between the affect of the covid921 to druggies population.

$H_A$ There is a difference between the affect of the covid921 test trials on druggies population.

```{r}
library(dplyr)
DrugUsers = filter(FinalData, FinalData$DrugUser == 'yes')
```

## Methods

Druggies is the categorical variable, Covid921 vaccine administration is the explanatory variable. Covid921s effectiveness is a response variable, therefore we utilize cat~cat helper file. 

## Descriptive Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")
```

This bar graph shows among the DrugUsers who received the vaccine had a higher percentage of getting Covid19. Compared to DrugUsers who received the placebo, those druggies had a lower chance of catching covid19. Those who never caught covid had little change. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

# Overall Results and Conclusions

